We begin by introducing causal attack trees.  This formulation of
attack trees was first proposed by Jhawar et al. \cite{Jhawar:2015}
where they called them causal attack trees, but sequential composition
does not always meet the same properties as conjunction, for example,
classically it is a self dual operator, thus, we follow Horne et al.'s
lead \cite{horne2017semantics} and call them causal attack trees.
\begin{definition}
  \label{def:atrees}
  Suppose $\mathsf{B}$ is a set of base attacks whose elements are
  denoted by $[[b]]$.  Then an \textbf{attack tree} is defined by
  the following grammar:
  \[
  \begin{array}{lll}
    [[A]],[[B]],[[C]],[[T]] := [[b]] \mid [[OR(A,B)]] \mid [[AND(A,B)]] \mid [[SEQ(A,B)]]\\
  \end{array}
  \]
  \noindent
  Equivalence of attack trees, denoted by $[[A ~ B]]$, is defined as
  follows:
  \begin{center}
    \begin{math}
      \setlength{\arraycolsep}{10px}
      \begin{array}{lll}
        [[OR(A,A) ~ A]]\\
        [[OR(OR(A,B),C) ~ OR(A,OR(B,C))]]\\
        [[AND(AND(A,B),C) ~ AND(A,AND(B,C))]]\\
        [[SEQ(SEQ(A,B),C) ~ SEQ(A,SEQ(B,C))]]\\
        [[OR(A,B) ~ OR(B,A)]]\\
        [[AND(A,B) ~ AND(B,A)]]\\
        [[AND(A,OR(B,C)) ~ OR(AND(A,B),AND(A,C))]]\\
        [[SEQ(A,OR(B,C)) ~ OR(SEQ(A,B),SEQ(A,C))]]\\
      \end{array}
    \end{math}
  \end{center}
\end{definition}
This definition of causal attack trees differs slightly from Jhawar
et. al.'s~\cite{Jhawar:2015} definition.  They define $n$-ary
operators, but we only consider the binary case, because it fits better
with the models presented here and it does not loose any generality
because we can model the $n$-ary case using binary operators in the
obvious way.

Throughout the sequel we will show that the previous rules are sound
with respect to our new models, but just as Horne et
al. \cite{horne2017semantics} we will then show that there are
properties of attack trees that these rules do not support.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: main.tex
%%% End: 
