\documentclass{llncs}

\usepackage{wrapfig}
\usepackage{amssymb,amsmath,mathtools}
\usepackage{cmll}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{mdframed}
\usepackage{thmtools, thm-restate}
\usepackage{enumitem}
\usepackage{minted}

%% This renames Barr's \to to \mto.  This allows us to use \to for imp
%% and \mto for a inline morphism.
\let\mto\to
\let\to\relax
\newcommand{\to}{\rightarrow}

% Commands that are useful for writing about type theory and programming language design.
%% \newcommand{\case}[4]{\text{case}\ #1\ \text{of}\ #2\text{.}#3\text{,}#2\text{.}#4}
\newcommand{\interp}[1]{\llbracket #1 \rrbracket}
\newcommand{\normto}[0]{\rightsquigarrow^{!}}
\newcommand{\join}[0]{\downarrow}
\newcommand{\redto}[0]{\rightsquigarrow}
\newcommand{\nat}[0]{\mathbb{N}}
\newcommand{\fun}[2]{\lambda #1.#2}
\newcommand{\CRI}[0]{\text{CR-Norm}}
\newcommand{\CRII}[0]{\text{CR-Pres}}
\newcommand{\CRIII}[0]{\text{CR-Prog}}
\newcommand{\subexp}[0]{\sqsubseteq}
%% Must include \usepackage{mathrsfs} for this to work.
\newcommand{\powerset}[0]{\mathscr{P}}

\date{}

% Cat commands.
\newcommand{\Four}[0]{\mathsf{4}}
\newcommand{\cat}[1]{\mathcal{#1}}
\newcommand{\catop}[1]{\cat{#1}^{\mathsf{op}}}
\newcommand{\Hom}[3]{\mathsf{Hom}_{\cat{#1}}(#2,#3)}
\newcommand{\limp}[0]{\multimap}
\newcommand{\dial}[1]{\mathsf{Dial_{#1}}(\mathsf{Sets})}
\newcommand{\dcSets}[1]{\mathsf{DC_{#1}}(\mathsf{Sets})}
\newcommand{\sets}[0]{\mathsf{Sets}}
\newcommand{\obj}[1]{\mathsf{Obj}(#1)}
\newcommand{\mor}[1]{\mathsf{Mor(#1)}}
\newcommand{\id}[0]{\mathsf{id}}
\newcommand{\lett}[0]{\mathsf{let}\,}
\newcommand{\inn}[0]{\,\mathsf{in}\,}
\newcommand{\cur}[1]{\mathsf{cur}(#1)}
\newcommand{\curi}[1]{\mathsf{cur}^{-1}(#1)}

\newenvironment{changemargin}[2]{%
  \begin{list}{}{%
    \setlength{\topsep}{0pt}%
    \setlength{\leftmargin}{#1}%
    \setlength{\rightmargin}{#2}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\itemindent}{\parindent}%
    \setlength{\parsep}{\parskip}%
  }%
  \item[]}{\end{list}}

%% Ott
\input{aterms-ott}

\begin{document}

\title{On Linear Logic, Functional Programming, and Attack Trees}

\author{Harley Eades III\inst{1} \and Jiaming Jiang\inst{2} \and Aubrey Bryant\inst{3}}

\institute{Computer Science, Augusta University, \href{mailto:harley.eades@gmail.com}{harley.eades@gmail.com}
\and
Computer Science, North Carolina State University
\and
Computer Science, Augusta University}

\maketitle

\begin{abstract}
  This paper has two main contributions. The first is a new linear
  logical semantics of causal attack trees in four-valued truth
  tables.  Our semantics is very simple and expressive, supporting
  specializations, and combines in an interesting way the \emph{ideal}
  and \emph{filter} semantics of causal attack trees. Our second
  contribution is Lina, a new embedded, in Haskell, domain specific
  functional programming language for conducting threat analysis using
  attack trees.  Lina has many benefits over existing tools; for
  example, Lina allows one to specify attack trees very abstractly,
  which provides the ability to develop libraries of attack trees,
  furthermore, Lina is compositional, allowing one to break down
  complex attack trees into smaller ones that can be reasoned about
  and analyzed incrementally.  Furthermore, Lina supports
  automatically proving properties of attack trees, such as
  equivalences and specializations, using Maude and the semantics
  introduced in this paper.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
\input{intro-output}

\vspace{-7px}
\section{Causal Attack Trees}
\label{sec:causal_attack_trees}
\input{attack-trees-output}
% section causal_attack_trees (end)

\vspace{-7px}
\section{A Quaternary Semantics for Causal Attack Trees}
\label{sec:a_quaternary_semantics_for_causal_attack_trees}
\input{quaternary-semantics-output}
% section a_ternary_semantics_for_causal_attack_trees (end)

\vspace{-7px}
\section{Lina: An EDSL for Conducting Threat Analysis using Causal Attack Trees}
\label{sec:lina:_an_edsl_for_conducting_threat_analysis_using_causal_attack_trees}
\input{Lina}
% section lina:_an_edsl_for_conducting_threat_analysis_using_causal_attack_trees (end)

\vspace{-7px}
\section{Conclusion and Future Work}
\label{sec:conclusion}
We made two main contributions: a new four-valued truth table
semantics of causal attack trees that supports specializations of
attack trees, and a new embedded domain specific programming language
called Lina for specifying, reasoning, and analyzing attack trees.

We plan to lift the quaternary semantics into a natural deduction
system based on the logic of bunched implications, and then study
proof search within this new system.  Lina is under active
development, and we have a number of extensions planned, for example,
adding support for attack-defense trees, attack(-defense) graphs,
attack nets, a GUI for viewing the various models, and a SMT backend.
Finally, it is necessary for number of case studies to be carried out
within Lina to be able to support the types of analysis required for
real world applications.  
% section conclusion (end)

\vspace{-7px}
\section{Acknowledgments}
\label{sec:acknowledgments}
This work was supported by NSF award \#1565557.  We thank Cl\'ement
Aubert for helpful discussions and feedback on previous drafts of this
paper.
% section acknowledgments (end)


\bibliographystyle{plain}
\bibliography{ref}

\appendix

\section{Maude Specification for Causal Attack Trees}
\label{sec:maude-spec-causal-attack-trees}
\begin{verbatim}
mod Causal is

protecting LOOP-MODE .

sorts Formula .
subsort Nat < Formula .

op _||_    : Formula Formula -> Formula [ctor assoc comm] .
op _._     : Formula Formula -> Formula [ctor assoc comm] .
op _;_     : Formula Formula -> Formula [ctor assoc] .
op EQ(_,_) : Formula Formula -> Bool .

var P Q R S : Formula .

eq P . (Q || R)  = (P . Q)  || (P . R) .
eq P ; (Q || R)  = (P ; Q)  || (P ; R) .
eq (Q || R) ; P  = (Q ; P)  || (R ; P) .

ceq EQ(P,Q) = true
  if P = Q .
eq EQ(P,Q) = false .

endm  
\end{verbatim}
% section  (end)


\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 

